Unimodular covers of 3-dimensional parallelepipeds and Cayley sums

### 2019-07-24 unimodular-covers.tex
last changes plus figures

\tableofcontents

@@

\subsection*{Acknowledgements:} We thank Akiyoshi Tsuchiya, Spencer Backman, and Johannes Hofscheier for posing these questions to us and 
Christian Haase for helpful discussions.

@@

The idea to have in mind is that of the regular tetrahedron inscribed in a cube. 
Modulo an affine transformation this is exactly the situation of $T$ and $C(T)$.
\giulia{we should add a picture of the standard cube}
\paco{perhaps, but not necessarily}

%\Giulia{not sure if we need this remark:
%$S$ is ``half a fundamental domain'' for the affine lattice $\Lambda_P$ generated by $p_1,\dots,p_4$, in the following precise sense: if $x$ is an arbitrary point, then $S$ intersects either $\Lambda_P + x$ or $\Lambda_P -x$ (and, if $x$ is generic then the intersection is unique). Proof: think of this property for the standard cube and the lattice $\{sum of coordinates is even\}$ generated by the vertices of the regular tetrahedron inscribed in it. (Remark: let us assume without loss of generality that $\Lambda_P$ contains the origin, so that $\Lambda + x$ is the class of $x$ modulo $\Lambda$).
%}

@@

\begin{proof}
By White's classification of empty tetrahedra (\cite{White1964}, see also, e.~g.~\cite[Sect.~4.1]{HPPS-survey}), there is no loss of generality in assuming $T=\conv(p_1,p_2,p_3,p_4)$ with
\[
p_1=(0,0,0), \quad
p_2=(1,0,0), \quad
p_3=(0,0,1), \quad
p_4=(a,b,1).
\]
where $b\ge 2$ is the (normalized) volume of $T$, and $a\in \{1,\dots,b-1\}$ satisfies $\gcd(a,b)=1$. This gives 
\paco{picture of the quadrilaterals $p_1q_4p_2q_3$ and $q_2p_3q_1p_4$ would be good.}
\paco{perhaps, but not necessarily}
\begin{align*}
q_1=\left(\frac{1+a}2,\frac{b}2,1\right), &&
q_2=\left(\frac{a-1}2,\frac{b}2,1\right), \quad\\
q_3=\left(\frac{1+a}2,\frac{b}2,0\right), &&
q_4=\left(\frac{1-a}2,-\frac{b}2,0\right).
\end{align*}

@@

Observe that $u+v=p_1+p_2=q_3+q_4$.
Now, the translation of vector $\left(\frac{a-1}2,\frac{b}2,1\right)$ sends the facet $\conv(p_1p_2q_3q_4)$ of $C(T)$ to the opposite facet $\conv(q_2q_1p_4p_3)$, so that the translated points $u + \left(\frac{a-1}2,\frac{b}2,1\right)$ and $v + \left(\frac{a-1}2,\frac{b}2,1\right)$ lie in $\conv(q_2q_1p_4p_3)$ and are centrally symmetrically placed in it. Thus, one of them belongs to $\conv(q_1p_4p_3) \subset T_2$ and the other to $\conv(q_2p_4p_3) \subset T_1$.
%\paco{Alternative arguments: one can explicitly compute coordinates of lattice points in $T_1$ and $T_2$, or one can say that the isomorphism $T_{a,b} \cong T_{a^{-1},b}$ of White tetrahedra exchanges the edges at $\{z=0\}$ and at $\{z=1\}$.}
%Now, this implies that the quadrilateral $\conv(p_1q_4p_2q_3)$ contains a fundamental domain for the lattice $\Z^2\times\{0\}$. Hence, its translate $\conv(q_2p_3q_1p_4)$ contains a fundamental domain for $\Z^2\times\{1\}$ and, in particular, it contains at least one lattice point other than $p_3$ and $p_4$. By central symmetry around its center $\left(\frac{a}2,-\frac{b}2,1\right)$, $\conv(q_2p_3q_1p_4)$ must contain lattice points in both triangles $\conv(q_2p_3p_4)\subset T_1$ and $\conv(q_1p_3p_4)\subset T_2$.
\end{proof}

@@

\paco{removed definitions of "nef cayley", "mixed IDP" and "weak Minkowski sum" from here}
%\begin{definition}
%In the above conditions, we say that $Q$ \emph{is a weak Minkowski summand of $P$} if the normal fan of $Q$ refines that of $P$. 
%If this happens, we say that $\cayley(P,Q)$ is a \emph{nef Cayley sum}.\giulia{I guess there's no nice term for this now...}
%\paco{not sure; maybe we can keep nef after all; another (unclear) option could be ``weak prism over $Q$'' or ``degenerate prism over $Q$''.}
%
%We say that the pair $(P,Q)$ is \emph{mixed IDP} if 
%\paco{``mixed IDP'' is not the right property, since this is weaker than IDP} \giulia{do we give it a name at all then?}
%\paco{perhaps mixed-IDP}
%\[
%(P\cap \Z^d) + (Q\cap \Z^d) = (P+Q)\cap \Z^d.
%\]
%\end{definition}
%
%In this language, \Cref{conj:Oda}\eqref{itm:mixedIDP} translates to:
%\emph{If $Q$ is nef for $P$ and $P$ is smooth then the pair $(P,Q)$ mixed IDP.}
%
The so-called \emph{Cayley Trick} is the isomorphism
\[
2\cayley(P, Q) \cap (\R^d\times \{1\}) \cong P+Q,
\]
which easily implies:

@@

We now turn our attention to $d=2$, in order to prove \Cref{thm:cayley}. A triangulation of $\cayley(P,Q)\subset \R^3$ consists of tetrahedra of types $(1,3)$, $(2,2)$ and $(3,1)$, where the type denotes how many vertices they have in $P$ and in $Q$. Empty tetrahedra of types $(1,3)$ or $(3,1)$, which are Cayley sums of a triangle in $P$ and a point in $Q$, or viceversa, are automatically unimodular. The case that we need to study are therefore tetrahedra of type $(2,2)$, which are Cayley sums of a segment $p\subset P$ and a segment $q\subset Q$.
%
%
%The unimodular prod-simplices are unimodular parallelograms $p+q$ with $p$ a segment of $P$ and $q$ a segment of $Q$, and unimodular triangles $T+u$, where $T \subset P$ and $u \in Q$ or viceversa. 
%
%Consider any lattice triangulation $\T$ of $\cayley(P,Q)$, or equivalently a mixed subdivision of $P+Q$. Observe that any simplex in $T$ has vertices in $P \times \{0\}$ or in $Q \times \{1\}$, since $\cayley(P,Q)$ lies between two consecutive lattice hyperplanes $\{x_3=0\}$ and $\{x_3=1\}$. Thus simplices in $\T$ are of three types; they can , or have three vertices in $P$ and one in $Q$ (type $(3,1)$), have one vertex in $P$ and three in $Q$ (type $(1,3)$) or have two vertices in $P$ and two in $Q$ (type $(2,2)$). 
%
\giulia{fix this}
\paco{done, but check it}
%
%As a consequence, we also have that:
%\begin{align*}
%\text{$\cayley(P,Q)$ has a unimodular cover} &\leftrightarrow \text{$P+Q$ can be covered by unimodular prod-simplices}.
%\end{align*}
The following lemma, whose proof we postpone to \Cref{sec:the_lemma}, is crucial to understand how to unimodularly cover these tetrahedra. 
%For this the following result, the proof of which we postpone to \Cref{sec:the_lemma}, is crucial. 
We use the following conventions: if $a, b$ are points, we denote by $[a,b]$ and $(a,b)$ respectively the closed and open segments with endpoints $a,b$. Given a segment $s=[a,b]$, we denote the vector $\vec s:= b-a$ and  the line spanned by $\vec s$ by $\vecline s$.

\begin{lemma}
\label{lemma:cayley}
%\paco{$Q$ is 2-dim, but $P$ may not}
Let $Q$ be a two-dimensional lattice polytope and $P$ a weak Minkowski summand of it. 
Let $p=[p_1,p_2] \subset P$ and $q=[q_1,q_2]\subset Q$ be two primitive and non-parallel lattice segments, and let $\vecline p$ and $ \vecline q$ be the lines spanned by them.  If the parallelogram $p + q$ is not unimodular, then at least one of the regions
\[
((p_1, p_2) + \vecline q ) \cap P, 
\qquad \text{and} \qquad
((q_1, q_2) + \vecline p ) \cap Q
\]
contains a lattice point. See \Cref{fig:strips}.
%\[
%(\int(p) + \vecline q) \cap P
%\qquad \text{or} \qquad
%\int(q + \vecline p) \cap Q
%\]
%contains a lattice point.
\end{lemma}
%\giulia{what I want is: a lattice point $p \in P$ such that either $p+q_1$ or $p+q_2$ lies in the interior of $\{p_1, p_2\} + \{q_1, q_2\}$ (or symmetric for $Q$)}

\begin{figure}[htb]
\scalebox{.75}{\input{strips.pdf_t}}
\caption{The strips  of Lemma \ref{lemma:cayley}}
\label{fig:strips}
\end{figure}

\begin{corollary}
\label{coro:covercayley}
\paco{inserted this corollary to make proof clearer, and to parallel section 2} \giulia{picture?}
Let $T$ be an empty lattice tetrahedron contained in the Cayley sum $\cayley(P,Q)$, where $P$ is a weak Minkowski summand of $Q$. Then, $T$ can be covered by unimodular tetrahedra contained in $\cayley(P,Q)$.
\end{corollary}

@@

Let us denote $\tilde u$, $\tilde p_1$, $\tilde p_2$, $\tilde q_1$, $\tilde q_2$ the points corresponding to $u, p_1, p_2, q_1, q_2$ in $\cayley(P,Q)$.
That is, $\tilde p_i = p\times\{0\}$, $\tilde q_i = p\times\{1\}$, and $\tilde u = u\times\{0\}$.
Observe that the assumption $u\in((p_1, p_2) + \vecline q$ implies that of the segments $[u,q_i]$ crosses the triangle $\conv(p_1,p_2,q_j)$, where $\{i,j\}=\{1,2\}$. In turn, this means that the polytope $\conv(\tilde u, \tilde p_1, \tilde p_2, \tilde q_1, \tilde q_2) = \cayley(t,q)$
 has the following two triangulations:
\begin{gather*}
\mathcal T^+:= \left\{ \cayley(p,q), \cayley(t, \{q_i\}) \right\},
\\
\mathcal T^-:= \{ \cayley([p_1,u],q), \cayley([p_2,u],q), \cayley(t, \{q_j\}) \}.
\end{gather*}
The tetrahedra $\cayley(t, \{q_j\})$ and $\cayley(t, \{q_i\})$ are unimodular, which implies that $T=\cayley(p,q)$ has volume equal to the sum of the volumes of $\cayley([p_1,u],q)$ and $\cayley([p_2,u],q)$. In particular, we have covered $T$ by the three tetrahedra in $\mathcal T^-$, which are of smaller volume and hence have unimodular covers by inductive assumption.
%
%A unimodular triangulation of a simplex of type $(3,1)$ (resp $(1,3)$) is obtained simply by unimodularly triangulating the triangular face lying in $P$ (resp $Q$) and coning over the vertex in $Q$ (resp $P$). This leaves simplices of type $(2,2)$. In the mixed subdivision of $P+Q$ corresponding to $T$, these are mixed cells, that is parallelograms with one edge in $P$ and one in $Q$ \giulia{how imprecise can be without being confusing?}.
%
%Let $S=\{p_1, p_2\} + \{q_1, q_2\}$ be such a parallelogram. By \Cref{lemma:cayley} we know that there is a lattice point $p \in P$ such that either $p+q_1$ or $p+q_2$ lies in $S$, w.l.o.g. suppose it is $p+q_1$. Then we can cover $S$ by the triangle $\{p_1, p_2, p\} + q_1$ and parallelograms $\{p_1, p\}  + \{q_1, q_2\}$ and $\{p_2, p\} + \{q_1, q_2\}$. As above, the triangle can be unimodularly triangulated, while the volume of the two parallelograms sums up to the volume of $S$, and therefore each has volume strictly smaller than $S$. By iterating this procedure, we can cover $S$ by unimodular cells.
\end{proof}

@@

Let $w=\area(p+q) \geq 2$, where $\area$ denotes the area normalized to a fundamental domain. Then 
\[
w=\width_{f_q}(p + \vecline q )=\width_{f_q}(p)=\width_{f_p}(q)=\width_{f_p}(q +\vecline p ). 
\] 
\paco{picture for setup needs change}

@@

\begin{proof}[Proof of \Cref{lemma:cayley}]
Suppose by contradiction that there is no lattice point as described in the lemma. In particular, no lattice point on the boundary of $P$ can be in the interior of the strip $p + \vecline q$.  Thus the boundary of $P$ contains two primitive segments which each have one vertex on each side of the strip $p + \vecline q$; we will call these $\ell=[l_1, l_2]$ and $r=[r_1, r_2]$, with $\ell$ and $r$ crossing the strip $p + \vecline q$  in $V_1$ and $ V_2$ respectively and the convention that $f_q(l_2) >f_q(l_1)$ and $f_q(r_2) >f_q(r_1)$. This readily implies 
\begin{gather}
\label{eq:widthp}
\begin{array}{cc}
f_q(l_1) \leq f_q(p_1), &
f_q(l_2) \geq f_q(p_2), \\
f_q(r_1) \leq f_q(p_1),&
f_q(r_2) \geq f_q(p_2).
\end{array}
\end{gather}

The same holds for $Q$ and the strip $q+\vecline p$, and we call the segments  $b=[b_1, b_2], t=[t_1, t_2]$ with $b$ and $t$ crossing the strip in $H_1$ and $H_2$ respectively. Again we have
\begin{gather}
\label{eq:widthq}
\begin{array}{cc}
f_p(t_1) \leq f_p(q_1), &
f_p(t_2) \geq f_p(q_2), \\
f_p(b_1) \leq f_p(q_1), &
f_p(b_2) \geq f_p(q_2).
\end{array}
\end{gather}

@@

Then $Q \cap V_2$ is contained in the open strip $\{-1<f_q(x)<w-1\}$, of width $w$. This cannot contain a translated copy of $r$, since $\width_{f_q}(r) \geq w$, see \Cref{fig:claim2} .
\end{proof}

@@

We assume w.l.o.g. that the maximum on $t$ (and hence on $b$) is achieved in $V_2$ \giulia{do we need it??}, that is to say, $f_p$ and $f_q$ increase in the same direction along $t$ (and hence along $b$).

@@

Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ on $d$ and $s_2$ on $d'$.  
Since $t$ separates $q_1$ and $q_1+q_2-p_1$ and $g$ decreases from $t_1$ to $t_2$ (by \Cref{claim:g}), the inequality $g(x)< g(d')$ holds on $Q\cap V_2$,
and in particular for $r_2'$. Since $r_2'$ is a lattice point, $g(r_2')\leq g(d)= g(r_1')$, which contradicts \Cref{claim:g}).
\end{proof}
\giulia{add a picture?}

>>>>

%\tableofcontents

@@

\subsection*{Acknowledgements:} We would like to thank Akiyoshi Tsuchiya, Spencer Backman, and Johannes Hofscheier for posing these questions to us and 
Christian Haase for helpful discussions.

@@

The idea to have in mind is that of the regular tetrahedron inscribed in a cube, see \Cref{fig:circumscribed_parall}. 
Modulo an affine transformation this is exactly the situation of $T$ and $C(T)$.

\begin{figure}[htb]
\includegraphics[width=.7\textwidth]{circumscribed_parall}
\caption{In red we have the tetrahedron $T$, in black the circumscribed parallelepiped $C(T)$, and in blue the corner simplex $T_4$.}
\label{fig:circumscribed_parall}
\end{figure}
\giulia{I added a picture, now we can decide if we want it in the paper or not.}

@@

\begin{proof}
By White's classification of empty tetrahedra (\cite{White1964}, see also, e.~g.~\cite[Sect.~4.1]{HPPS-survey}), there is no loss of generality in assuming $T=\conv(p_1,p_2,p_3,p_4)$ with
\[
p_1=(0,0,0), \quad
p_2=(1,0,0), \quad
p_3=(0,0,1), \quad
p_4=(a,b,1).
\]
where $b\ge 2$ is the (normalized) volume of $T$, and $a\in \{1,\dots,b-1\}$ satisfies $\gcd(a,b)=1$. This gives 
\begin{align*}
q_1=\left(\frac{1+a}2,\frac{b}2,1\right), &&
q_2=\left(\frac{a-1}2,\frac{b}2,1\right), \quad\\
q_3=\left(\frac{1+a}2,\frac{b}2,0\right), &&
q_4=\left(\frac{1-a}2,-\frac{b}2,0\right).
\end{align*}

@@

See \Cref{fig:corner}. Observe that $u+v=p_1+p_2=q_3+q_4$.
Now, this implies that the quadrilateral $\conv(p_1q_4p_2q_3)$ contains a fundamental domain for the lattice $\Z^2\times\{0\}$. Hence, its translate $\conv(q_2p_3q_1p_4)$ contains a fundamental domain for $\Z^2\times\{1\}$ and, in particular, it contains at least one lattice point other than $p_3$ and $p_4$. By central symmetry around its center $\left(\frac{a}2,-\frac{b}2,1\right)$, $\conv(q_2p_3q_1p_4)$ must contain lattice points in both triangles $\conv(q_2p_3p_4)\subset T_1$ and $\conv(q_1p_3p_4)\subset T_2$.

\begin{figure}[htb]
\includegraphics[width=.9\textwidth]{Whites}
\caption{The parallelograms $\conv(p_1p_2q_3q_4)$ and $\conv(q_2q_1p_4p_3)$. The red points are lattice points.}
\label{fig:corner}
\end{figure}
\giulia{ditto}

%\paco{Alternative arguments: one can explicitly compute coordinates of lattice points in $T_1$ and $T_2$, or one can say that the isomorphism $T_{a,b} \cong T_{a^{-1},b}$ of White tetrahedra exchanges the edges at $\{z=0\}$ and at $\{z=1\}$.}

\end{proof}

@@

The so-called \emph{Cayley Trick} is the isomorphism
\[
2\cayley(P, Q) \cap (\R^d\times \{1\}) \cong P+Q,
\]
which easily implies:

@@

We now turn our attention to $d=2$, in order to prove \Cref{thm:cayley}. A triangulation of $\cayley(P,Q)\subset \R^3$ consists of tetrahedra of types $(1,3)$, $(2,2)$ and $(3,1)$, where the type denotes how many vertices they have in $P$ and in $Q$. Empty tetrahedra of types $(1,3)$ or $(3,1)$, which are Cayley sums of a triangle in $P$ and a point in $Q$, or viceversa, are automatically unimodular. The case that we need to study are therefore tetrahedra of type $(2,2)$, which are Cayley sums of a segment $p\subset P$ and a segment $q\subset Q$.
\giulia{fix this}
\paco{done, but check it}
\giulia{just changed small things}
The following lemma, whose proof we postpone to \Cref{sec:the_lemma}, is crucial to understand how to unimodularly cover these tetrahedra. 
We use the following conventions: if $a, b$ are points, we denote by $[a,b]$ and $(a,b)$ respectively the closed and open segments with endpoints $a,b$. Given a segment $s=[a,b]$, we denote the vector $\vec s:= b-a$ and  the line spanned by $\vec s$ by $\vecline s$.

\begin{lemma}
\label{lemma:cayley}
%\paco{$Q$ is 2-dim, but $P$ may not}
Let $Q$ be a two-dimensional lattice polytope and $P$ a weak Minkowski summand of it. 
Let $p=[p_1,p_2] \subset P$ and $q=[q_1,q_2]\subset Q$ be two primitive and non-parallel lattice segments, and let $\vecline p$ and $ \vecline q$ be the lines spanned by them.  If the parallelogram $p + q$ is not unimodular, then at least one of the regions
\[
((p_1, p_2) + \vecline q ) \cap P, 
\qquad \text{and} \qquad
((q_1, q_2) + \vecline p ) \cap Q
\]
contains a lattice point. See \Cref{fig:strips}.
%\[
%(\int(p) + \vecline q) \cap P
%\qquad \text{or} \qquad
%\int(q + \vecline p) \cap Q
%\]
%contains a lattice point.
\end{lemma}

\begin{figure}[htb]
\scalebox{.75}{\input{strips.pdf_t}}
\caption{The strips  of Lemma \ref{lemma:cayley}}
\label{fig:strips}
\end{figure}
\giulia{added picture, we may not want it}

\begin{corollary}
\label{coro:covercayley}
Let $T$ be an empty lattice tetrahedron contained in the Cayley sum $\cayley(P,Q)$, where $P$ is a weak Minkowski summand of $Q$. Then, $T$ can be covered by unimodular tetrahedra contained in $\cayley(P,Q)$.
\end{corollary}

@@

Let us denote $\tilde u$, $\tilde p_1$, $\tilde p_2$, $\tilde q_1$, $\tilde q_2$ the points corresponding to $u, p_1, p_2, q_1, q_2$ in $\cayley(P,Q)$.
That is, $\tilde p_i = p\times\{0\}$, $\tilde q_i = p\times\{1\}$, and $\tilde u = u\times\{0\}$.
Observe that the assumption $u\in((p_1, p_2) + \vecline q$ implies that of the segments $[u,q_i]$ crosses the triangle $\conv(p_1,p_2,q_j)$, where $\{i,j\}=\{1,2\}$, see \Cref{fig:flip}.

\begin{figure}[htb]
\includegraphics[width=.9\textwidth]{flip}
\caption{$[u,q_2]$ intersects $\conv(p_1,p_2,q_1)$}
\label{fig:flip}
\end{figure}

In turn, this means that the polytope $\conv(\tilde u, \tilde p_1, \tilde p_2, \tilde q_1, \tilde q_2) = \cayley(t,q)$
 has the following two triangulations:
\begin{gather*}
\mathcal T^+:= \left\{ \cayley(p,q), \cayley(t, \{q_i\}) \right\},
\\
\mathcal T^-:= \{ \cayley([p_1,u],q), \cayley([p_2,u],q), \cayley(t, \{q_j\}) \}.
\end{gather*}
The tetrahedra $\cayley(t, \{q_j\})$ and $\cayley(t, \{q_i\})$ are unimodular, which implies that $T=\cayley(p,q)$ has volume equal to the sum of the volumes of $\cayley([p_1,u],q)$ and $\cayley([p_2,u],q)$. In particular, we have covered $T$ by the three tetrahedra in $\mathcal T^-$, which are of smaller volume and hence have unimodular covers by inductive assumption.
%
%A unimodular triangulation of a simplex of type $(3,1)$ (resp $(1,3)$) is obtained simply by unimodularly triangulating the triangular face lying in $P$ (resp $Q$) and coning over the vertex in $Q$ (resp $P$). This leaves simplices of type $(2,2)$. In the mixed subdivision of $P+Q$ corresponding to $T$, these are mixed cells, that is parallelograms with one edge in $P$ and one in $Q$ \giulia{how imprecise can be without being confusing?}.
%
%Let $S=\{p_1, p_2\} + \{q_1, q_2\}$ be such a parallelogram. By \Cref{lemma:cayley} we know that there is a lattice point $p \in P$ such that either $p+q_1$ or $p+q_2$ lies in $S$, w.l.o.g. suppose it is $p+q_1$. Then we can cover $S$ by the triangle $\{p_1, p_2, p\} + q_1$ and parallelograms $\{p_1, p\}  + \{q_1, q_2\}$ and $\{p_2, p\} + \{q_1, q_2\}$. As above, the triangle can be unimodularly triangulated, while the volume of the two parallelograms sums up to the volume of $S$, and therefore each has volume strictly smaller than $S$. By iterating this procedure, we can cover $S$ by unimodular cells.
\end{proof}

@@

Let $w=\area(p+q) \geq 2$, where $\area$ denotes the area normalized to a fundamental domain. Then:
\[
w=\width_{f_q}(p + \vecline q )=\width_{f_q}(p)=\width_{f_p}(q)=\width_{f_p}(q +\vecline p ). 
\]

@@

\begin{proof}[Proof of \Cref{lemma:cayley}]
Suppose by contradiction that there is no lattice point as described in the lemma. In particular, no lattice point on the boundary of $Q$ can be in the interior of the strip $q + \vecline p$.  Thus the boundary of $Q$ contains two primitive segments which each have one vertex on each side of the strip $q + \vecline p$; we will call these  $b=[b_1, b_2], t=[t_1, t_2]$, with $b$ and $t$ crossing the strip in $H_1$ and $H_2$ respectively and the convention that $f_p(b_2) >f_p(b_1)$ and $f_p(t_2) >f_p(t_1)$. This readily implies

\begin{gather}
\label{eq:widthq}
\begin{array}{cc}
f_p(t_1) \leq f_p(q_1), &
f_p(t_2) \geq f_p(q_2), \\
f_p(b_1) \leq f_p(q_1), &
f_p(b_2) \geq f_p(q_2).
\end{array}
\end{gather}

The same holds for $P$ and the strip $p+\vecline q$, and we call the segments $\ell=[l_1, l_2]$ and $r=[r_1, r_2]$, with $\ell$ and $r$ crossing the strip $p + \vecline q$  in $V_1$ and $ V_2$ respectively. The only difference is that in the case that $P$ is one dimensional we have $\ell=r=p$.  Again we have
\begin{gather}
\label{eq:widthp}
\begin{array}{cc}
f_q(l_1) \leq f_q(p_1), &
f_q(l_2) \geq f_q(p_2), \\
f_q(r_1) \leq f_q(p_1),&
f_q(r_2) \geq f_q(p_2).
\end{array}
\end{gather}

@@

Then $Q \cap V_2$ is contained in the open strip $\{-1<f_q(x)<w-1\}$, of width $w$. This cannot contain a translated copy of $r$, since $\width_{f_q}(r) \geq w$, see \Cref{fig:claim2}. This is a contradiction, since $P$ is a Minkowski summand of $Q$ and therefore $Q$ must have an edge parallel to $t$.
\end{proof}

@@

We assume w.l.o.g. that the maximum on $t$ (and hence on $b$) is achieved in $V_2$, that is to say, $f_p$ and $f_q$ increase in the same direction along $t$ (and hence along $b$).

@@

Consider the segment $s$ contained in $r_1' + \vecline p$ with endpoints $s_1=r_1'$ on $d$ and $s_2$ on $d'$.  
Since $t$ separates $q_1$ and $q_1+q_2-p_1$ and $g$ decreases from $t_1$ to $t_2$ (by \Cref{claim:g}), the inequality $g(x)< g(d')$ holds on $Q\cap V_2$,
and in particular for $r_2'$. Since $r_2'$ is a lattice point, $g(r_2')\leq g(d)= g(r_1')$, which contradicts \Cref{claim:g}).
\end{proof}
